---
title: "qPRAentry workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{qPRAentry workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(qPRAentry)
```

# Introduction

`qPRAentry` is a package for...

Two application examples of the functions available in the `qPRAentry` package are 
shown below. The example [A](#example-ISO) uses the functions applicable to any country in the world using ISO codes. The example [B](#example-NUTS) uses the functions designed for use with the NUTS code system.

In both cases, trade data are required for the commodity that is considered to be a potential pathway for the pest under assessment. The data required include:

  - Import of the commodity from third countries to the countries of interest in 
  the study, both the total quantity and the quantity imported from third countries 
  where the pest is known to be present. 
  - Trade of the commodity between the countries of interest.
  - Internal production of the commodity in the countries of interest.

# A. Example using ISO codes {#example-ISO}

## 1. Trade data preparation

### Trade data required

CSV data files can be loaded via `load_csv()`

```{r}
data("datatrade_NorthAm")
```

Total quantity of commodity imported from third countries. Using the example data, 
we select imports from all third countries (column partner):

```{r}
extra_total <- datatrade_NorthAm$extra_import
head(extra_total)
```

Quantity of commodity imported from third countries where the pest is present. In 
this case we assume that the pest is present in countries ``NTR_1" and ``CNTR_2":

```{r message=FALSE, warning=FALSE}
library(dplyr)
CNTR_pest <- c("CNTR_1", "CNTR_2")
extra_pest <- datatrade_NorthAm$extra_import %>% filter(partner%in%CNTR_pest)
head(extra_pest)
```

Quantity of commodity traded between countries of interest:

```{r}
intra_trade  <- datatrade_NorthAm$intra_trade
head(intra_trade)
```

Quantity of commodity produced internally in each country of interest:

```{r}
internal_production  <- datatrade_NorthAm$internal_production
head(internal_production)
```

### Generation of the `TradeData` object from data frames

```{r}
trade_NorthAm <- trade_data(extra_total = extra_total,
                            extra_pest = extra_pest,
                            intra_trade = intra_trade,
                            internal_production = internal_production)
```

Total trade:

```{r}
head(trade_NorthAm$total_trade)
```

Trade between countries:

```{r}
head(trade_NorthAm$intra_trade)
```

### Plot

Plot the total available quantity of commodity available in each country

```{r}
library(ggplot2)
plot_countries(data = trade_NorthAm$total_trade,
               iso_col = "country_IDs", 
               values_col = "total_available") +
  xlim(-180,-20) + ylim(0,90)
```

## 2. Ntrade

### Calculation of the Ntrade for each time period

```{r}
ntrade_NorthAm <- ntrade(trade_data = trade_NorthAm)
head(ntrade_NorthAm)
```

### Ntrade summary for the time periods

```{r}
ntrade_NorthAm_summary <- ntrade(trade_data = trade_NorthAm,
                                 summarise_result = c("mean", "sd", 
                                                      "quantile(0.025)", 
                                                      "median",
                                                      "quantile(0.975)"))
head(ntrade_NorthAm_summary)
```

### Plot 

```{r}
# Plot the median of Ntrade
library(ggplot2)
plot_countries(data = ntrade_NorthAm_summary,
               iso_col = "country_IDs", 
               values_col = "median") +
  xlim(-180,-20) + ylim(0,90)
```


## 3. Ntrade redistribution

```{r}
# ## Example of data redistribution in Northern American countries
# data(datatrade_NorthAm)
# # extract ISO codes at country level
# iso_3166_1 <- unique(datatrade_NorthAm$internal_production$reporter)
# # simulate values for each country
# iso1_data <- data.frame(iso_3166_1 = iso_3166_1,
#                         values = abs(rnorm(length(iso_3166_1), 30000, 10000)))
# # simulate values for each subdivision
# # ISO 3166_2 codes extracted from 'rnaturalearth' package
# library(dplyr)
# library(rnaturalearth)
# iso2_data <- ne_states() %>% 
#   # subdivisions of the countries selected above
#   filter(substr(iso_3166_2, 1, 2) %in% iso_3166_1) %>% 
#   distinct(iso_3166_2) %>%
#   select(iso_3166_2) %>% 
#   # simulate values for each subdivision
#   mutate(values = abs(rnorm(nrow(.), 0, 1000)))
# 
# data_redist <- redist_iso(data = iso1_data,
#                           iso_col = "iso_3166_1",
#                           values_col = "values",
#                           redist_data = iso2_data,
#                           redist_iso_col = "iso_3166_2",
#                           redist_values_col = "values")
# 
# head(data_redist)
# 
# # Plot using the 'sf' object from 'rnaturalearth' package
# iso2_plot <- ne_states() %>% 
#   filter(substr(iso_3166_2, 1, 2) %in% iso_3166_1) %>% 
#   left_join(data_redist, by = join_by(iso_3166_2 == ISO_2))
# 
# library(ggplot2)
# ggplot(iso2_plot, aes(fill = values)) + 
#   geom_sf()
#   
# ## Example of redistribution of two columns
# # simulate values for each country
# iso1_data$values2 <- abs(rnorm(nrow(iso1_data), 1000, 500))
# data_redist <- redist_iso(data = iso1_data,
#                           iso_col = "iso_3166_1",
#                           values_col = c("values", "values2"),
#                           redist_data = iso2_data,
#                           redist_iso_col = "iso_3166_2",
#                           redist_values_col = "values")
# 
# head(data_redist)
```


## 4. Pathway model

```{r}
# ## Example using Northern American countries and ntrade simulated data
# data("datatrade_NorthAm")
# # Extract country IDs and simulate ntrade data
# IDs <- datatrade_NorthAm$internal_production$reporter
# df <- data.frame(IDs = IDs,
#                  ntrade_values = abs(rnorm(length(IDs), 10000, 2000)))
# # Expression for the pathway model using 3 parameters
# eq <- "(1/P1)*P2*P3"
# # Distribution for each parameter
# parameters <- list(
#   P1 = list(dist = "beta", shape1 = 0.5, shape2 = 1),
#   P2 = list(dist = "gamma", shape = 1.5, scale = 100),
#   P3 = list(dist = "lnorm", mean = 5, sd = 2)
# )
# # Run pathway_model()
# res_pathway <- pathway_model(ntrade_data = df,
#                              IDs_col = "IDs",
#                              values_col = "ntrade_values",
#                              expression = eq,
#                              parameters = parameters,
#                              niter = 100)
# head(res_pathway)
# # summary of the total for all countries
# res_pathway[res_pathway$IDs == "Total",]
# # plot
# plot_countries(res_pathway, "IDs", "Median")
```

# B. Example using NUTS codes {#example-NUTS}

## 1. Trade data preparation

```{r}
## Example with simulated trade data for Europe 
# with selected countries and a specific time period
# Load data
data("datatrade_EU")
# Total extra-import data: the total import is identified as partner "Extra_Total"
extra_total <- datatrade_EU$extra_import %>% filter(partner=="Extra_Total")
# Extra-import data from countries where the pest is present
extra_pest <- datatrade_EU$extra_import %>% filter(partner!="Extra_Total")
# Intra-trade data
intra_trade  <- datatrade_EU$intra_trade
# Internal production data
internal_production  <- datatrade_EU$internal_production
# Sample 5 countries from data
filter_IDs <- sample(unique(extra_total$reporter), 5)
# Generate trade data (TradeData object)
trade_EU <- trade_data(extra_total = extra_total,
                       extra_pest = extra_pest,
                       intra_trade = intra_trade,
                       internal_production = internal_production,
                       filter_IDs = filter_IDs,
                       filter_period = 2020)
# Plot the total available quantity of commodity available in each country
plot_countries(data = trade_EU$total_trade, 
               iso_col = "country_IDs", 
               values_col = "total_available") +
  xlim(-30,50) + ylim(25,70)
```


## 2. Ntrade


```{r}
## Example with simulated trade data for Europe 
# Load data
data("datatrade_EU")
# Total extra-import data: the total import is identified as partner "Extra_Total"
extra_total <- datatrade_EU$extra_import %>% filter(partner=="Extra_Total")
# Extra-import data from countries where the pest is present
extra_pest <- datatrade_EU$extra_import %>% filter(partner!="Extra_Total")
# Intra-trade data
intra_trade  <- datatrade_EU$intra_trade
# Internal production data
internal_production  <- datatrade_EU$internal_production
# Generate trade data (TradeData object)
trade_EU <- trade_data(extra_total = extra_total,
                       extra_pest = extra_pest,
                       intra_trade = intra_trade,
                       internal_production = internal_production)
# Ntrade mean and sd for the time periods
ntrade_EU <- ntrade(trade_data = trade_EU,
                    summarise_result = c("mean", "sd"))
# Plot Ntrade mean
plot_countries(data = ntrade_EU, 
               iso_col="country_IDs", 
               values_col="mean") +
  xlim(-40,50) + ylim(25,70)
# Ntrade for selected countries and a specific time period
# Sample 5 countries from trade data
country_IDs <- sample(unique(trade_EU$total_trade$country_IDs), 5)
ntrade_EU_s <- ntrade(trade_data = trade_EU,
                      filter_IDs = country_IDs,
                      filter_period = 2020)
head(ntrade_EU_s)
# Plot Ntrade result
plot_countries(data = ntrade_EU_s, 
               iso_col="country_IDs", 
               values_col="Ntrade_2020") +
  xlim(-40,50) + ylim(25,70)

```


## 3. Ntrade redistribution

```{r}
# 
# #' ## Example of data redistribution to NUTS2 using population data
# data("datatrade_EU")
# # extract NUTS0 codes (country level)
# nuts0 <- unique(datatrade_EU$internal_production$reporter)
# # simulate values for each country
# nuts0_data <- data.frame(nuts0 = nuts0,
#                          values = abs(rnorm(length(nuts0), 30000, 10000)))
# 
# data_redist <- redist_nuts(data = nuts0_data,
#                            nuts_col = "nuts0",
#                            values_col = "values",
#                            to_nuts = 2,
#                            redist_data = "population",
#                            population_year = c(2017, 2018, 2019))
# head(data_redist)
# 
# # Plot
# plot_nuts(data = data_redist,
#           nuts_level = 2,
#           nuts_col = "NUTS2", 
#           values_col = "values")
# 
# ## Example of data redistribution of two value columns to NUTS1 using custom data
# nuts0_data$values2 <- abs(rnorm(length(nuts0), 2000, 500))
# # NUTS1 codes extracted from 'giscoR' package
# library(dplyr)
# library(giscoR)
# nuts1_data <- gisco_get_nuts(nuts_level=1) %>% 
#   select(NUTS_ID) %>% 
#   # simulate values for each NUTS1
#   mutate(values = abs(rnorm(nrow(.), 0, 1000)))
# 
# data_redist <- redist_nuts(data = nuts0_data,
#                            nuts_col = "nuts0",
#                            values_col = c("values", "values2"),
#                            to_nuts = 1,
#                            redist_data = nuts1_data,
#                            redist_nuts_col = "NUTS_ID",
#                            redist_values_col = "values")
# 
# head(data_redist)
# 
# # Plot
# plot_nuts(data = data_redist,
#           nuts_level = 1,
#           nuts_col = "NUTS1", 
#           values_col = "values2")
```


## 4. Pathway model

```{r}
# ## Example using Northern American countries and ntrade simulated data
# data("datatrade_NorthAm")
# # Extract country IDs and simulate ntrade data
# IDs <- datatrade_NorthAm$internal_production$reporter
# df <- data.frame(IDs = IDs,
#                  ntrade_values = abs(rnorm(length(IDs), 10000, 2000)))
# # Expression for the pathway model using 3 parameters
# eq <- "(1/P1)*P2*P3"
# # Distribution for each parameter
# parameters <- list(
#   P1 = list(dist = "beta", shape1 = 0.5, shape2 = 1),
#   P2 = list(dist = "gamma", shape = 1.5, scale = 100),
#   P3 = list(dist = "lnorm", mean = 5, sd = 2)
# )
# # Run pathway_model()
# res_pathway <- pathway_model(ntrade_data = df,
#                              IDs_col = "IDs",
#                              values_col = "ntrade_values",
#                              expression = eq,
#                              parameters = parameters,
#                              niter = 100)
# head(res_pathway)
# # summary of the total for all countries
# res_pathway[res_pathway$IDs == "Total",]
# # plot
# plot_countries(res_pathway, "IDs", "Median")
```



