---
title: "Ntrade report"
output: pdf_document
params:
  time_period: NA
  units: NA
  Nt_result: NA
  Nt_redist: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,
                      fig.pos = 'h',
                      fig.align = 'center',
                      fig.height = 3,
                      fig.width = 4)
```

```{r}
time_period <- params$time_period
units <- params$units
Nt <- params$Nt_result
Nt_redist<- params$Nt_redist
```


#Introduction

This report presents the results of the $N_{trade}$ analysis using the Shiny application. Data was processed using the *qPRAentry* package.

$N_{trade}$ is defined as the total quantity of commodities (infested by a pest or not) imported by a country (e.g., an EU country) from other third-countries where the pest is present. This application is meant to calculate $N_{trade}$ and redistribute the data into the countries considered.


```{r}
time_period <- params$time_period
units <- params$units
Nt <- params$Nt_result
Nt_redist<- params$Nt_redist
```

*$N_{trade}$ reuslts*

Below, the $N_{trade}$ results are displayed in both tabular and map formats.
The map shows the median $N_{trade}$ values by country at the NUTS0 level.

```{r}
summary_table <- Nt %>%
  group_by(NUTS0) %>%
  summarize(
    Mean = mean(Median, na.rm = TRUE),
    SD = sd(Median, na.rm = TRUE),
    Min = min(Median, na.rm = TRUE),
    Max = max(Median, na.rm = TRUE)
  )

# Display the table
kable(summary_table, caption = "$N_{trade}$ results by Country") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


plot_countries(Nt, "NUTS0", "Median", title = "Ntrade Median") +
   xlim(-20, 40) + ylim(35,70) +
   theme_bw()

```
Figure 1: Median $N_{trade}$ values by country at the NUTS0 level.


## $N_{trade}$ redistribution results

The $N_{trade}$ values can  be redistributed proportionally to NUTS2 regions based on population data or user-defined data.
Below, the $N_{trade}$ redistribution results are displayed in both tabular and map formats.
The map shows the median $N_{trade}$ values by regions (NUTS2 level).

