---
title: "Pathway model report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
bibliography: references.bib
params:
  ntrade: NA
  model_def: NA
  param_names: NA
  par_settings: NA
  n_iter: NA
  dist_result: NA
  Ninf: NA
  Ninf_EU: NA
  nuts_level: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,
                      fig.pos = 'h',
                      fig.align = 'center',
                      fig.height = 3,
                      fig.width = 7)
```

```{r}
ntrade <- params$ntrade # Ntrade data
model_def <- params$model_def # model equation
param_names <- params$param_names # parameter names
n_iter <- params$n_iter # number of iterations
par_settings <- params$par_settings # Parameter distributions
dist_result <- params$dist_result # random values generated from the distributions
Ninf <- params$Ninf # Ninf results
Ninf_EU <- params$Ninf_EU # Ninf results total EU
NUTS_lev <- params$nuts_level #selected NUTS (depending on the Excel Ntrade file chosen)

```

```{r}
# function to print the model equation on multiple lines
split_equation <- function(eq, width = 70) {
  parts <- strsplit(eq, "(?<=\\*|\\/|\\+|\\-)", perl = TRUE)[[1]]
  output <- ""
  line_length <- 0
  for (part in parts) {
    line_length <- line_length + nchar(part)
     if (grepl("\\*", part)) {
      part <- gsub("\\*", " \\\\cdot ", part)
    }
    if (line_length > width) {
      output <- paste0(output, part, " \\\\ ")
      line_length <- nchar(part)
    } else {
      output <- paste0(output, part, " ")
    }
  }
  return(output)
}

# width: width of the line
model_def_split <- split_equation(model_def, width=100)
```


## Introduction

This report details the results of the pathway model analysis conducted using 
the Pathway model app of the \texttt{qPRAentry} package.
It provides an estimation of the number of potential founder populations 
$(\mathit{NPFP})$ of a pest in different regions, using $N_{trade}$ data combined 
with additional user-defined parameters.

The European Food Safety Authority (EFSA) utilises pathway models to evaluate the 
risk of pests entering the EU territories. The complexity of the pathway model 
can vary based on factors such as the specific pest under consideration and the 
availability of detailed data.

The pathway model combines trade data on potentially infested commodities with 
other parameters that are involved in the risk of pest entry, from the source country of 
origin to the EU. Additional information can be found in the EFSA Guidance on 
quantitative pest risk assessment [@efsa2018guidance].



## Pathway model

The defined pathway model and its corresponding parameters are presented below:

\begin{gather*}
`r model_def_split`
\end{gather*}


A total of `r n_iter` iterations were performed to randomly generate parameter 
values based on the specified distribution.

### Parameters: 


```{r, results='asis'}
tex_to_expr <- function(tex_str) {
  tex_str <- gsub("\\$", "", tex_str) 
  tex_str <- gsub("_\\{([^}]+)\\}", "[\\1]", tex_str) 
  tex_str <- gsub("_([[:alnum:]])", "[\\1]", tex_str)
  tex_str <- gsub("\\^\\{([^}]+)\\}", "^\\1", tex_str)
  tex_str <- gsub("\\^([[:alnum:]])", "^\\1", tex_str)

  parse(text = tex_str)[[1]]
}

par(mfrow = c(1, 2), mar = c(4, 3, 2, 1))

for (i in seq_along(param_names)) {
  par_name <- tex_to_expr(param_names[i])
  # Histogram
  hist(dist_result[[i]], 
       main = bquote(.(par_name) ~ "~" ~ .(par_settings[i])),
       xlab = "",
       probability = TRUE, 
       breaks = "fd", 
       col = "lightblue",
       cex.main = 0.8,
       cex.lab = 0.8) 
  # Add a density line
  lines(density(dist_result[[i]]), col = "blue", lwd = 2)
}

```


## $\mathit{NPFP}$ results

As a result of the pathway model the number of potential founder populations $(\mathit{NPFP})$ 
is estimated for each NUTS `r NUTS_lev` level.
The summary of the overall results for all the selected NUTS are shown below.

```{r}

# Display the table
knitr::kable(Ninf_EU, caption = "NPFP results for all the selected NUTS", digits = 4) 

```

The map below shows the median $\mathit{NPFP}$ values by NUTS`r NUTS_lev`.

```{r, fig.width=8, fig.height=8}

plot_nuts(Ninf, "NUTS_ID", "Median", nuts_level = NUTS_lev,
          title = "NPFP Median", legend_title="NPFP") +
   xlim(-20, 40) + ylim(35,70) +
   theme_bw()+
  theme(
     plot.title = element_text(size = 14),
     legend.title = element_text(size = 12)
   )

```
\begin{center}
Figure 1: $\mathit{NPFP}$ at the NUTS`r NUTS_lev` level.
\end{center}


## References


